---
title: "Monthly export predictions"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, error = F, warning = F, echo = FALSE)
library(tidyverse);library(lubridate);library(zoo);library(plotly);library(coffeestats)
coffeestats::loadMTS()
coffeestats::loadScaleExports()
coffeestats::setDataDir()

# Set months you're working with. 
checkMonth <- max(mts$month)
checkMonthText <- format(checkMonth, "%B %Y")
predMonth  <- max(mts$month) + months(1)
predMonthText <- format(predMonth, "%B %Y")
```

## Introduction

This is an html document which summarises the scaleExports function, and automatically produces a report. 

It should be produced once a month, after the new ICO export data is available. 

It also summarises the findings from last month, and compares them to the actual data which is now available. 

This report will check the results for `r checkMonthText`, and offer a prediction for `r predMonthText`.  

### Monthly Trade Statistics

```{r, include = FALSE}we
# Recreate last month's scaleExports.
exports2 <- full_join(
    x = mts %>% filter(month != checkMonth),
    y = readr::read_csv(file.path(coffeestats, "cropyears.csv")) %>% 
        rename(country = Country, cyGroup = Crop.year)
) %>% 
    na.omit() %>% 
    arrange(country, month) %>% 
    mutate(cropYear = ifelse(month.abb[lubridate::month(month)] == cyGroup,
                             lubridate::year(month),
                             NA)) %>% 
    tidyr::fill(cropYear) %>% 
    filter(cropYear <= lubridate::year(month)) %>% 
    na.omit()
exportsAv2 <- exports2 %>% 
    group_by(country, cropYear) %>% 
    filter(n() == 12) %>% 
    arrange(country, month) %>% 
    mutate(totalExports = sum(value),
           share = value/totalExports) %>% 
    group_by(country, lubridate::month(month)) %>% 
    mutate(avShare = mean(share)) %>% 
    ungroup() %>% 
    select(country, month = `lubridate::month(month)`, avShare) %>% 
    distinct() %>% 
    group_by(country) %>% 
    mutate(cumSum = cumsum(avShare))
checkPrediction <- full_join(
    x = exports2 %>% 
        group_by(country, cropYear) %>% 
        summarise(total = sum(value, na.rm=TRUE)) %>% 
        filter(cropYear == max(cropYear)),
    y = exportsAv2 %>% filter(month == lubridate::month(checkMonth - months(1))),
    by = "country"
) %>% 
    mutate(scaleExports = total/cumSum) %>% 
    select(country, cropYear, scaleExports) %>% 
    full_join(
        x = .,
        y = filter(exportsAv2, month == month(checkMonth))
    ) %>% 
    mutate(predExports = scaleExports * avShare, month = checkMonth) %>% 
    select(country, cropYear, predMonth = month, scaleExports, predExports) %>% 
    na.omit()

full_join(
    checkPrediction %>% 
        select(country, cropYear, oldScale = scaleExports),
    scaleExports %>% 
        select(country, cropYear, newScale = scaleExports)
) %>% 
    mutate(diff = newScale - oldScale) %>% 
    readr::write_csv(file.path(coffeestats, paste0(lubridate::today(), "-checkScale.csv")))
```

> Summary of what happened in `r checkMonthText`

Total exports in `r checkMonthText` came to `r mts %>% filter(month == checkMonth, country == "TOTAL") %>% magrittr::extract2('value') %>% format(., big.mark = ",")` bags, compared to `r mts %>% filter(month == checkMonth - months(1), country == "TOTAL") %>% magrittr::extract2('value') %>% format(., big.mark = ",")` in `r format(checkMonth - months(1), "%B %Y")`. 

```{r, include = FALSE}
# Get a comparison of prediction against actual 
comparison <- full_join(
    x = mts %>% filter(month == checkMonth, country != "TOTAL", !grepl("Arabica|Robusta", country)),
    y = checkPrediction %>% select(country, month = predMonth, prediction = predExports)
) %>% 
    mutate(diff = value - prediction) %>% 
    arrange(desc(abs(diff)))
```

Compared to the prediction, there were major differences in `r paste0(head(comparison$country, 5), collapse = ", ")`, see the graph below. 

```{r, echo = FALSE, out.width="100%"}
library(plotly)
compPlot <- comparison %>% 
    slice(1:20) %>% 
    mutate(country = ordered(country, levels = country))
p1 <- plot_ly(compPlot,
        x = ~diff, y = ~country, type = "bar", orientation = "h", name = "Difference",
        marker = list(color = "#619fb5"),
        text = ~paste0(country, ": ", round(diff, 1)), hoverinfo = "text"
    ) %>% 
    layout(
        title = paste("Actual vs. predicted exports for", checkMonthText),
        xaxis = list(tickangle = 325, title = "", zeroline = FALSE),
        yaxis = list(title = "", separatethousands = TRUE, autorange = "reversed", side = "right")
    )
p2 <- plot_ly(compPlot, x = ~country) %>% 
    add_bars(
        y = ~prediction, name = "Prediction",
        marker = list(color = "#ef8066"),
        text = ~paste(country, "prediction:", round(prediction, 1)), hoverinfo = "text") %>% 
    add_bars(
        y = ~value, name = "Actual",
        marker = list(color = "#ffcb90"),
        text = ~paste(country, "actual:", round(value, 1)), hoverinfo = "text") %>% 
    layout(xaxis = list(title = "", tickangle = 325),
           yaxis = list(title = "", zeroline = FALSE),
           barmode = "group")
subplot(p2, p1) %>% 
    layout(
        legend = list(orientation = "h", xanchor = "center", yanchor = "top", y = -0.15, x = 0.6),
        margin = list(r = 120)
    )
```


### scaleExports

> Again, this is a prediction for `r predMonthText`

Total exports for `r predMonthText` are estimated at `r format(round(sum(scaleExports$predExports, na.rm = TRUE), 1), big.mark = ",")` bags, compared to `r format(round(sum((scaleExportsRaw %>% filter(month == predMonth - years(1)))$value, na.rm = TRUE), 1), big.mark = ",")` for `r format(predMonth - years(1), "%B %Y")`. 

```{r, echo = FALSE, out.width = "100%"}
# Like predCountry graph, but for everyone
# Going to just use Oct-Sep for everyone, rather than crop years
coffeeYear <- ordered(c(10:12, 1:9),
                      levels = c(10:12, 1:9))
globalExports <- exportsRaw %>% 
    group_by(month) %>% 
    summarise(totValue = sum(value, na.rm = TRUE)) %>% 
    mutate(cropYear = ifelse(month(month) == 10, year(month), NA)) %>% 
    mutate(cropYear = zoo::na.locf(cropYear, na.rm = FALSE)) %>% 
    na.omit()
globalExportsAv <- globalExports %>% 
    group_by(cropYear) %>% 
    mutate(n = n()) %>% 
    filter(n == 12) %>% 
    arrange(month) %>% 
    mutate(totalExports = sum(totValue),
           share = totValue/totalExports) %>% 
    group_by(month(month)) %>% 
    mutate(avShare = mean(share)) %>% 
    ungroup() %>% 
    select(month = `month(month)`, avShare) %>% 
    distinct() %>% 
    mutate(cumSum = cumsum(avShare))
globalExportsPred <- full_join(
    x = globalExports %>%
        top_n(n = 1, wt = cropYear) %>% 
        mutate(month = month(month),
               totalYear = sum(totValue)),
    y = globalExportsAv
) %>% 
    mutate(scaleExports = totalYear/cumSum,
           scaleExports = min(scaleExports, na.rm = TRUE),
           predExports = scaleExports * avShare) %>%
    select(month, actual = totValue, prediction = predExports, scaleExports) %>% 
    full_join(
        x = .,
        y = globalExports %>% 
            group_by(month(month)) %>%
            summarise(min = min(totValue, na.rm = TRUE),
                      max = max(totValue, na.rm = TRUE)) %>% 
            rename(month = `month(month)`)
    ) %>% 
    mutate(month = ordered(month, 
                           labels = month.abb[c(10:12, 1:9)],
                           levels = month))
p1 <- plot_ly(globalExportsPred, x = ~month) %>% 
    add_lines(name = "Prediction", y = ~prediction,
              line = list(color = "#cc593d", dash = "dash"),
              text = ~paste0(month, " prediction: ", format(round(prediction, 0), big.mark = ",")),
              hoverinfo = "text") %>% 
    add_lines(name = "Actual", y = ~actual, 
              line = list(color = "#cc593d"),
              text = ~paste0(month, " actual: ", format(round(actual, 0), big.mark = ",")),
              hoverinfo = "text") %>% 
    add_ribbons(name = "Min/max", line = list(width = 0), 
                ymin = ~min, ymax = ~max, hoverinfo = "none",
                fillcolor = "#ffe090", opacity = 0.4) %>% 
    layout(title = "Export prediction function for current coffee year",
           xaxis = list(title = "", tickangle = 270),
           yaxis = list(title = "", separatethousands = TRUE, ticklen = 20, tickcolor = "#FFF"),
           margin = list(r = 40))
p2 <- bind_rows(
    globalExports %>% 
        group_by(cropYear) %>% 
        mutate(n = n()) %>% 
        filter(n == 12) %>% 
        summarise(total = sum(totValue)),
    data_frame(cropYear = max(globalExports$cropYear),
               total = max(globalExportsPred$scaleExports))
) %>% 
    plot_ly(., x = ~cropYear, y = ~total, type = "bar", 
            marker = list(color = "#6a9eb5"),
            text = ~paste0(cropYear, ": ", format(round(total, 0), big.mark = ",")), name = "Total",
            hoverinfo = "text") %>% 
    layout(xaxis = list(title = ""),
           yaxis = list(title = "", zeroline = FALSE))
subplot(p1, p2, widths = c(0.6, 0.4), nrows = 1, margin = 0.05) %>% 
    layout(legend = list(orientation = "h", xanchor = "center", yanchor = "top", y = -0.2, x = 0.5))
```

## Appendix

Full scaleExports table for `r predMonthText`. 

```{r, echo = FALSE}
knitr::kable(scaleExports %>% 
                 na.omit() %>% 
                 mutate(cropYear = as.character(cropYear)) %>% 
                 filter(predExports >= 1, scaleExports >= 0) %>% 
                 select(country, cropYear, predExports, scaleExports), 
             digits = 0,
             row.names=FALSE,
             col.names = c("Country", "Crop year", predMonthText, "Full year"),
             align = c("lcrr"),
             caption = "",
             format.args = list(big.mark = ","))
```

